---
title: "Hands-on_Ex03"
author: "Andy Tan"
date: "27 August 2024"
date-modified: "last-modified"
output:
  html-documnet:
    toc: yes
---

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```
# 3.1 Importing the spatial data
```{r}
childcare_sf <- st_read("/Users/andy/Desktop/IS415-VAA/Hands-on_Ex/Hands-On_Ex03/data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```
```{r}
sg_sf <- st_read(dsn = "/Users/andy/Desktop/IS415-VAA/Hands-on_Ex/Hands-On_Ex03/data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "/Users/andy/Desktop/IS415-VAA/Hands-on_Ex/Hands-On_Ex03/data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

# 3.2 Mapping Geospatial data points
```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```
tmap set to plot

# 3.3 Geospatial Data wrangling
```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
childcare
```

```{r}
mpsz
```

```{r}
sg
```

# 3.4 Converting the Spatial* class into generic sp format
```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```

# 3.5 Converting the generic sp format into spatstat’s ppp format
```{r}
childcare_ppp <- as.ppp(st_coordinates(childcare_sf), st_bbox(childcare_sf))
```

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

```{r}
any(duplicated(childcare_ppp))
```

```{r}
multiplicity(childcare_ppp)
```

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

```{r}
tmap_mode('view')
```

```{r}
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
any(duplicated(childcare_ppp_jit))
```
Spot darker spots

# 3.6 Creating owin object
```{r}
sg_owin <- as.owin(sg_sf)
```

```{r}
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

# 3.7 Combining point events object and owin object
```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
summary(childcareSG_ppp)
```

# 3.8 Kernel Density Estimation
```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
```


```{r}
plot(kde_childcareSG_bw)
```

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

# 3.9 Recalling KDE value
```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
```

```{r}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```
# 3.10 Working with different automatic badwidth methods
```{r}
bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```


```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

# 3.11 Working with different kernel methods
```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```
# 3.12 Computing KDE by using fixed bandwidth
```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```
# 3.13 Computing KDE by using adaptive bandwidth
```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```
# 3.14 Converting KDE output into grid object.
```{r}
kde_childcareSG.im <- as.im(kde_childcareSG.bw)
kde_childcareSG.im
```

```{r}
gridded_kde_childcareSG_bw <- as(kde_childcareSG.im, "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_bw)
```
# 3.15 Converting gridded output to raster
```{r}
kde_childcareSG_bw_raster <- raster(kde_childcareSG.bw)
kde_childcareSG_bw_raster
```
# 3.16 Assigning Projection System
```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```
# 3.17 Visualising output in tmap
```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("layer", palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
# 3.18 Comparing Spatial Points using KDE
```{r}
pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")

#plotting the target areas
par(mfrow=c(2,2))
plot(pg, main = "PUNGGOL")
```

```{r}
plot(tm, main = "TAMPINES")
```

```{r}
plot(ck, main = "CHOA CHU KANG")
```
```{r}
plot(jw, main = "JURONG WEST")
```
# 3.19 Creating owin object
```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

# 3.20 Childcare points and study area
```{r}
# number of childcare points within the area
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]

# converting metre to km
childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")

# plot
par(mfrow=c(2, 2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

```{r}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```
# 3.21 Nearest neighbour analysis
Testing spatial point patterns using Clark and Evans Test
```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```
Clark and Evans Test: Choa Chu Kang planning area
```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
Clark and Evans Test: Tampines planning area
```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```